Alone Together: Compositional Reasoning and Inference for Weak Isolation

### 2017-07-08 reasoning.tex
Edits to Sec. 2.

\rulelabel{RG-Txn} is the top-level rule that lets us prove a
transaction preserves the high-level invariant $I$ when executed under
the required isolation as specified by $\I$. It relies on the
transaction-local judgment to verify the body ($c$) of a transaction
$i$. The precondition $P$ of $c$ must follow from the fact that the
transaction-local database ($\stl$) is initially empty, and the global
database satisfies the high-level invariant $I$. The rely relation
($\R_l$) for the local judgment is a ternary relation computed as $R$
modulo $\I_e$:
\begin{smathpar}
\begin{array}{lcl}
%\R_l(\stl,\stg,\stg') & \Leftrightarrow & R(\stg,\stg') \conj
%\I_e(\stl,\stg,\stg')
  \R_l(\stl,\stg,\stg') & \Leftrightarrow & \exists \stg_1. \I_e(\stl, \stg_1, \stg) \wedge R(\stg, \stg') \wedge \I_e(\stl, \stg_1, \stg')
\end{array}
\end{smathpar}
In words, if there exists $\stg_1$ which was the global database state during the last step by a transaction and if the transaction can take its next step at the current global database state $\stg$, then even after the interference, the transaction should still be able to take its next step. If there is no such state $\stg_1$, then a transaction will never be able to take its next step after the interference, and hence such interference can be safely discarded.

@@

\begin{definition}[\bfseries Semantics of RG judgments]
\label{def:rg-semantics}
The semantics of the global and transaction-local RG judgments are
defined thus:
\begin{smathpar}
\begin{array}{lcl}
\R \vdash \hoare{P}{c}{Q} & \defeq & \forall \stl,\stl',\stg,\stg'.~
  P(\stl,\stg) \conj (\tbox{c}_i,\stl,\stg) \rstepssto{n}
  (\tbox{\cskip}_i, \stl',\stg')
  \Rightarrow Q(\stl',\stg')\\
\rg{I,R}{c}{G,I} & \defeq &  \forall \stg.\, I(\stg)
  \Rightarrow (\forall n,\stg'.\; (c,\stg) 
    \rstepssto{n} (\cskip,\stg') \Rightarrow I(\stg')) \\
&&\hspace*{0.5in}\conj \texttt{step-guaranteed}(R,G,c,\stg)\\
\end{array}
\end{smathpar}

>>>>

\rulelabel{RG-Txn} is the top-level rule that lets us prove a
transaction preserves the high-level invariant $I$ when executed under
the required isolation as specified by $\I$. It relies on the
transaction-local judgment to verify the body ($c$) of a transaction
$i$. The precondition $P$ of $c$ must follow from the fact that the
transaction-local database ($\stl$) is initially empty, and the global
database satisfies the high-level invariant $I$. The rely relation
($\R_l$) for the local judgment is a ternary relation computed as $R$
modulo $\I_e$:
\begin{smathpar}
\begin{array}{lcl}
%\R_l(\stl,\stg,\stg') & \Leftrightarrow & R(\stg,\stg') \conj
%\I_e(\stl,\stg,\stg')
  \R_l(\stl,\stg,\stg') & \Leftrightarrow & \exists \stg_1. \I_e(\stl, \stg_1, \stg) \wedge R(\stg, \stg') \wedge \I_e(\stl, \stg_1, \stg')
\end{array}
\end{smathpar}
In words, if there exists $\stg_1$ which was the global database state during the last step by a transaction and if the transaction can take its next step at the current global database state $\stg$, then even after the interference, the transaction should still be able to take its next step. If there is no such state $\stg_1$, then a transaction will never be able to take its next step after the interference, and hence such interference can be safely discarded. Often, this significantly simplifies the form of $\R_l$ irrespective of $R$. For example, when a transaction is executed against a snapshot of the database (i.e. $\mathbb{I}_{ss}$), $\R_l$ will be the identity function, since if $\stg \neq \stg'$, there does not exist any $\stg_1$ which satisfies the constraint for $\R_l(\stl,\stg,\stg')$.

@@

\begin{definition}[\bfseries Semantics of RG judgments]
\label{def:rg-semantics}
The semantics of the global and transaction-local RG judgments are
defined thus:
\begin{smathpar}
\begin{array}{lclr}
\R \vdash \hoare{P}{c}{Q} & \defeq & \forall \stl,\stl',\stg,\stg',
  \bar{v}.~ P(\stl,\stg) \conj (\tbox{c}_i,\stl,\stg) \rstepssto{n}
  (\tbox{\cskip}_i, \stl',\stg')
  \Rightarrow Q(\stl',\stg') & \bar{v} = {\sf freevars}(c)\\
\rg{I,R}{c}{G,I} & \defeq &  \forall \stg.\, I(\stg)
  \Rightarrow (\forall \stg'.\; (c,\stg) 
    \rstepssto{n} (\cskip,\stg') \Rightarrow I(\stg')) \\
&&\hspace*{0.5in}\conj \texttt{step-guaranteed}(R,G,c,\stg)\\
\end{array}
\end{smathpar}
